*Ctrl+Shift+Enter* - run code
*Ctrl+Alt+I* - insert new chunk
*install.packages("PackageName")* - install library


```{r}
# load the packages
library(RWeka)
library(caTools)
library(dplyr)
library("Rgraphviz")
library(partykit)
library(caret)
library(ROCR)

```

Load IRIS dataset
```{r}
set.seed(17)  # set a seed to get replicable results

iris <- read.csv(file="datasets/iris.csv", head=FALSE, sep=";", fileEncoding="UTF-8-BOM")
iris <- iris[sample(1:nrow(iris)), ]  # shuffle

dim(iris)
# plot(iris)  # show dataset
```

Split dataset to parts in a simple way: train - 80% and test - 20%
```{r}
spl_size <- floor(0.8 * nrow(iris))
set.seed(17)  # set a seed to get replicable results
train_ind <- sample(seq_len(nrow(iris)), size = spl_size)

train <- iris[train_ind, ]
test <- iris[-train_ind, ]

head(test, 5)

```

Train model
```{r}
model <- J48(V5 ~ V1+V2+V3+V4, data=train)
model

```

Visualize model
```{r}
ff <- tempfile()
write_to_dot(model, ff)
plot(agread(ff))
plot(model)

```

Get predictions about test data
```{r}
predictions <- predict(model, test[,1:4])  # make predictions
# table(predictions, test$V5) 
confusion_matrix <- confusionMatrix(predictions, test$V5, mode = "prec_recall")
confusion_matrix

```

Try to get some statistics out of confusion matrix
```{r}
# attributes(confusion_matrix)
# confusion_matrix$byClass
# class(confusion_matrix$byClass)

# str(confusion_matrix)
matrix_by_class <- confusion_matrix$byClass
matrix_by_class[, 'Precision']

precision_for_3_class <- unname(matrix_by_class[, 'Precision'])[3]

```

Get metrics for IRIS dataset
```{r}
stats <- confusion_matrix$byClass

acc <- mean(unname(stats[, 'Balanced Accuracy']))
prec <- mean(unname(stats[, 'Precision']))
rec <- mean(unname(stats[, 'Recall']))
fsc <- mean(unname(stats[, 'F1']))

cat("Statistics for IRIS:", acc, prec, rec, fsc)
```


Trying training model with cross validation
```{r}
folds_indices <- createFolds(iris$V5, k=10)
model <- train(V5 ~ V1+V2+V3+V4, method="J48", data=iris, tuneLength = 5, 
               trControl = trainControl(method="cv", indexOut=folds_indices))
```

```{r}
# ff <- tempfile()
# write_to_dot(model, ff)
# plot(agread(ff))
# plot(model)

model
plot(model)
plot(model$finalModel)

```

*C4.5-like Trees*
method = 'J48'
Type: Classification

Tuning parameters:
C (Confidence Threshold)
M (Minimum Instances Per Leaf)

C - The confidence is used to compute a pessimistic upper bound on the error rate at a leaf/node. The smaller this value, the more pessimistic the estimated error is and generally the heavier the pruning. J48 does not accept values greater the 0.5. 
